---
title: "Computational Musicology Storyboard"
date: "2025-02-26"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: cerulean
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}

library(tidyverse) 
library(gridExtra) 
library(rjson) 
library(ggplot2)
source("compmus.R")
knitr::opts_chunk$set(echo = FALSE)
```

## Analyzing my AI-generated MP3-files

Hier komt een klein inleidend stukje over mijn files, hoe ik ze heb
gegenereerd, welke prompts ik heb gebruikt en wat mijn verwachtingen
zijn.

## Plots

### Chromagram using "wietske-b-1.json" file and ceptogram

```{r fig.height=25, fig.width=35}
chroma1 <- "features/wietske-b-1.json" |>                           # Change the track
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()   # Change the theme?

chroma2 <- "features/wietske-b-2.json" |>                           # Change the track
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()   # Change the theme?

grid.arrange(chroma1, chroma2, ncol = 2)
```

### My ceptogram

Column {data-width=500}

```{r}
cepto1 <- "features/wietske-b-1.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                      # Change the theme?

cepto2 <- "features/wietske-b-2.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL
  ) +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic() # Change the theme?

grid.arrange(cepto1, cepto2, ncol = 2)
```

### My chroma-based self-similarity

```{r}
selfc1 <- "features/wietske-b-1.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

selfc2 <- "features/wietske-b-2.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(option = "plasma", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

grid.arrange(selfc1, selfc2, ncol = 2)
```

### My timbre-based self-similarity change pls what

```{r}
selft1 <- "features/wietske-b-1.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(option = "inferno", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

selft2 <- "features/wietske-b-2.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(option = "inferno", guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?

grid.arrange(selft1, selft2, ncol = 2)
```

### My chordogram

```{r}
library("tibble")

# C C# D Eb E F F# G Ab A Bb B

major_chord <- 
  c( 1, 0, 0, 0, 1, 0, 0, 1, 0, 0, 0, 0) 
minor_chord <-
  c( 1, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0) 
seventh_chord <- 
  c( 1, 0, 0, 0, 1, 0, 0, 1, 0, 0, 1, 0)

major_key <- 
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88) 
minor_key <- 
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <- 
  tribble( 
    ~name, ~template, 
    "Gb:7", circshift(seventh_chord, 6), 
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10), 
    "Db:maj", circshift(major_chord, 1), 
    "F:min", circshift(minor_chord, 5), 
    "Ab:7", circshift(seventh_chord, 8), 
    "Ab:maj", circshift(major_chord, 8), 
    "C:min", circshift(minor_chord, 0), 
    "Eb:7", circshift(seventh_chord, 3), 
    "Eb:maj", circshift(major_chord, 3), 
    "G:min", circshift(minor_chord, 7), 
    "Bb:7", circshift(seventh_chord, 10), 
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2), 
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5), 
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0), 
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4), 
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7), 
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2), 
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6), 
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9), 
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4), 
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8), 
    "B:7", circshift(seventh_chord, 11), 
    "B:maj", circshift(major_chord, 11), 
    "D#:min",circshift(minor_chord, 3) 
  )

key_templates <- tribble( ~name, ~template, "Gb:maj",
circshift(major_key, 6), "Bb:min", circshift(minor_key, 10), "Db:maj",
circshift(major_key, 1), "F:min", circshift(minor_key, 5), "Ab:maj",
circshift(major_key, 8), "C:min", circshift(minor_key, 0), "Eb:maj",
circshift(major_key, 3), "G:min", circshift(minor_key, 7), "Bb:maj",
circshift(major_key, 10), "D:min", circshift(minor_key, 2), "F:maj",
circshift(major_key, 5), "A:min", circshift(minor_key, 9), "C:maj",
circshift(major_key, 0), "E:min", circshift(minor_key, 4), "G:maj",
circshift(major_key, 7), "B:min", circshift(minor_key, 11), "D:maj",
circshift(major_key, 2), "F#:min", circshift(minor_key, 6), "A:maj",
circshift(major_key, 9), "C#:min", circshift(minor_key, 1), "E:maj",
circshift(major_key, 4), "G#:min", circshift(minor_key, 8), "B:maj",
circshift(major_key, 11), "D#:min", circshift(minor_key, 3) )

chord1 <- "features/wietske-b-1.json" |> 
compmus_chroma(norm = "manhattan") |>
compmus_match_pitch_templates(
  chord_templates, 
  norm = "identity",
  distance = "cosine"
)|> 

ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() + 
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = "Template", fill = NULL) + 
  theme_classic() 
# Try different norms (and match it with what you used in `compmus_chroma`)
# Try different distance metrics ) 


chord2 <- "features/wietske-b-2.json" |> 
compmus_chroma(norm = "manhattan") |>
compmus_match_pitch_templates(
  chord_templates, 
  norm = "identity",
  distance = "cosine"
)|> 

ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() + 
  scale_fill_viridis_c(guide = "none") + 
  labs(x = "Time (s)", y = "Template", fill = NULL) + 
  theme_classic() 
# Try different norms (and match it with what you used in `compmus_chroma`)
# Try different distance metrics ) 

grid.arrange(chord1, chord2, ncol = 2)
```

### Novelty functions: Energy

```{r}
nove1 <- "features/wietske-b-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")

nove2 <- "features/wietske-b-2.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")

grid.arrange(nove1, nove2, ncol = 2)
```

### Novelty functions: Spectral

```{r}
novs1 <- "features/wietske-b-1.json" |>
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty")

novs2 <- "features/wietske-b-2.json" |>
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty")

grid.arrange(novs1, novs2, ncol = 2)
```

### Tempograms: cyclic is false snap je

```{r}
tempo1f <- "features/wietske-b-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

tempo2f <- "features/wietske-b-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

grid.arrange(tempo1f, tempo2f, ncol = 2)
```

### Tempograms: cyclic is true

```{r} 
tempo1t <- "features/wietske-b-1.json" |> compmus_tempogram(window_size
= 8, hop_size = 1, cyclic = TRUE) |> ggplot(aes(x = time, y = bpm,
fill = power)) + geom_raster() + scale_fill_viridis_c(guide = "none") +
labs(x = "Time (s)", y = "Tempo (BPM)") + theme_classic()

tempo2t <- "features/wietske-b-2.json" |> compmus_tempogram(window_size
= 8, hop_size = 1, cyclic = TRUE) |> ggplot(aes(x = time, y = bpm,
fill = power)) + geom_raster() + scale_fill_viridis_c(guide = "none") +
labs(x = "Time (s)", y = "Tempo (BPM)") + theme_classic()

grid.arrange(tempo1t, tempo2t, ncol = 2)
```

### Week 12: Pre-processing
```{r}
library(tidymodels)
library(ggdendro)
library(heatmaply)

source("compmus.R")

get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  

compmus2025 <- read_csv("compmus2025.csv")

cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus_dist <- dist(cluster_juice, method = "euclidean")
```
### Hierarchical clustering
```{r}

compmus_dist |> 
  hclust(method = "single") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
```
### Heatmap
```{r}
heatmaply(
  cluster_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)
```

### Classification
```{r}
compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))

classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())      # Converts to z-scores.
  # step_range(all_predictors())    # Sets range to [0, 1].

compmus_cv <- compmus2025_filtered |> vfold_cv(5)

knn_model <-
  nearest_neighbor(neighbors = 1) |>
  set_mode("classification") |> 
  set_engine("kknn")

classification_knn <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(knn_model) |> 
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))

classification_knn |> get_conf_mat()

classification_knn |> get_conf_mat() |> autoplot(type = "mosaic")

classification_knn |> get_conf_mat() |> autoplot(type = "heatmap")

classification_knn |> get_pr()


```


### Random Forest
```{r}
forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")

indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit_resamples(
    compmus_cv, 
    control = control_resamples(save_pred = TRUE)
  )

indie_forest |> get_pr()

workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit(compmus2025_filtered) |> 
  pluck("fit", "fit", "fit") |>
  ranger::importance() |> 
  enframe() |> 
  mutate(name = fct_reorder(name, value)) |> 
  ggplot(aes(name, value)) + 
  geom_col() + 
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Importance")

compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Valence",
    y = "Arousal",
    size = "Tempo",
    colour = "AI"
  )


```
